---
title: "R Notebook"
---

# STAT 131A, Fall 2024: Homework 1 Coding Questions

## Question 1

This question continues where Lab 1 concluded.

Make sure to run the cell below before getting started.

```{r}
# Load in additional functions
library(tidyverse)
library(lubridate)
library(glue)

# Use three digits past the decimal point
options(digits = 3)

# Format plots with a white background and dark features.
theme_set(theme_bw())

# This is where the data is stored on DataHub.
STOPS_PATH = "../lab01/data/nyc_stops.rds"

# Read in the data
stops = read_rds(STOPS_PATH)

# Create population data for 2010
pop_2010 = tribble(
  ~race, ~pop,
  # Black, non-Hispanic
  "Black", 1776891,
  # Hispanic, any race
  "Hispanic", 2490350,
  # White, non-Hispanic
  "White", 2719856
)

pop_2010 = pop_2010 %>% 
  mutate(
    prop = pop/sum(pop)
  )

```

#### Part A

Calculate the proportion of all stops, frisks, and weapon recoveries that occurred among NYC individuals of each race and ethnicity.

- For example, your output data frame should provide the proportion of all frisks conducted in NYC that were of Black individuals. This proportion should be a number between 0 and 1.

- Your output data frame should have three rows.

```{r}
# Your code here!

# As a reminder, here's how we did this for stops:
stop_counts = stops %>%
  group_by(race) %>%
  summarize(n_stops = n()) %>%
  ungroup()

```

What do you conclude from your findings? How could someone use these results to mislead their audience?

- Answer in no more than seven sentences. 

- Write your answer as a code comment.

```{r}
# Your answer as a code comment here!

```

#### Part B

For each race and ethnicity, calculate the ratio of (a) stop proportion, (b) frisk
proportion, and (c) weapon proportion (calculated in Part A) to the corresponding 
population proportion.

- For example, your output dataframe should provide the ratio of (1) the proportion of all stops involving a Black individual and (2) the proportion
of the NYC residential population that identifies as Black.

- In total, you should calculate three ratios for each race and ethnicity. 

- The idea here is to assess how many more or fewer stops, frisks, and weapon recoveries we observe relative to each group's share of the population.

```{r}
# Your code here!

```

What do you conclude from your findings? How could someone use these results to mislead their audience?

- Answer in no more than seven sentences.

- Write your answer as a code comment.

```{r}
# Your answer as a code comment here!

```


#### Part C

First, filter the `stops` data to individuals who identify as Black. 

Second, for Black individuals in each precinct, calculate the fraction of stops that resulted in a frisk (the "frisk rate"). 

Store your result in a dataframe.

Repeat the two steps above for white individuals.

You should have two dataframes in total. Print them both.

```{r}
# Your code here!


```

#### Part D

For each precinct, calculate the ratio of the frisk rate of Black individuals to the frisk rate of white individuals. 

- This will require you to merge the two dataframes that you created in Part C.

```{r}
# Your code here!

```

What do you conclude from your findings? How could you present this result concisely to a non-technical audience?

- Answer in no more than five sentences.

```{r}
# Your answer as a code comment here!

```

#### Part E

Using your merged data from Part D, create a plot with the following features:

- On the x-axis, plot the frisk rate for Black individuals.
- On the y-axis, plot the frisk rate for white individuals.
- Size the points by the number of stops of Black and white individuals in each precinct.
- Add a dashed "y=x" line to your plot.

```{r}
# Your code here!

```

What do you conclude from your findings?

- Answer in no more than five sentences.

```{r}
# Your answer as a code comment here!

```

#### Part F

Repeat Part E, but instead of calculating frisk rates, calculate the proportion of **frisks** (not stops) that resulted in a weapon recovery ("hit rate").

```{r}
# Your code here!

```

What do you conclude from your findings?

- Answer in no more than seven sentences.

```{r}
# Your answer as a code comment here!

```

#### Concluding remarks on Question 1

In the final exercise, you conducted an **outcome test**. 

- Figure 3 in [this paper](https://5harad.com/papers/100M-stops.pdf) shows the result of an outcome test applied to large-scale policing data. 

Outcome tests can be used to detect potential discrimination not just in policing.
For example, you might think about how you could use an outcome test to detect discrimination in bank lending or hiring decisions.

## Question 2

Drawing on the summary statistics and descriptors from lecture (e.g., symmetry, modality, and skewness), describe the distribution of starting salaries of a representative sample of law school graduates from 2023.
The data is in `data/law-school-salaries.csv`.

- Support your data description by making one *relative* frequency histogram, one boxplot, and one paired boxplot. (You might take a moment to think about how much you learned from your summary statistics versus your plots!)

- Your data description should be no more than five sentences. 

- Write about what you think is most important! At the minimum, you should give some sense of the data distribution's shape, center, and spread.

- You are welcome to make additional plots if you'd like, but only these three are required. 

- Recall the difference between an EDA plot (your eyes only!) and a plot that is meant to communicate information to others.
Your plots should have the formatting of the latter, not the former. 

*Note: This problem uses a synthetic, but representative, sample of salaries based on publicly reported aggregate data. The actual salaries of individual law school graduates are confidential.*

```{r}
# Your code for importing the data and calculating summary statistics here!

```

```{r}
# Your histogram code here!

```

```{r}
# Your boxplot code here!

```

```{r}
# Your paired boxplot code here!

```

```{r}
# Your data description as a code comment in this cell!

```

Do have a guess as to why the distribution of law school new grad salaries looks the way it does?

- Answer in no more than three sentences. Write your answer as a code comment.

- Note that this answer is graded on thoughtfulness, not on correctness. You are welcome to Google the topic to learn more. 

```{r}
# Write your answer as a code comment in this cell!


```